Towards Practical First-Order Model Counting

### 2025-07-07 doc/talk/talk.tex
Finish explaining the algorithm

\documentclass{beamer}
\usepackage[utf8]{inputenc}
\usepackage[UKenglish]{babel}
\usepackage[UKenglish]{isodate}
\usepackage{tikz}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{complexity}
\usepackage{fontawesome}
\usepackage{dashbox}

@@

  \begin{tikzcd}
    g(0, m) \arrow[d, gray, visible on=<2->] & \phantom{-} \arrow[d, phantom, visible on=<2->, "\text{\structure{4.1. ($\Rightarrow$)} Construct the corr.\ sentence}"] & g(l, 0) \arrow[d, gray, visible on=<2->] \\
    |[visible on=<2->]| \forall x \in \Gamma\text{. }S(x) \lor \neg S(x) \arrow[d, gray, visible on=<3->] & \phantom{-} \arrow[d, phantom, visible on=<3->, "\text{\structure{4.2.} Recurse}"] & |[visible on=<2->]| \top \arrow[d, gray, visible on=<3->] \\
    |[visible on=<3->]| g(0, m) = \alt<4->{0^m}{\text{???}} \arrow[rd, gray, visible on=<5->] & |[visible on=<5->]| \text{\structure{4.3.} Add to }\structure{\mathcal{E}} & |[visible on=<3->]| g(l, 0) = \alt<4->{1}{???} \arrow[ld, gray, visible on=<5->] \\
    & |[visible on=<5->]| \mathcal{E} \arrow[d, gray, visible on=<6->, "\text{\color{black}{\structure{5.} Compile to C++}}"] & \\
    |[visible on=<7->]| \text{Domain sizes} \arrow[r, gray, visible on=<7->] & |[visible on=<6->]| \text{\dbox{C++ code}} \arrow[r, gray, visible on=<7->] & |[visible on=<7->]| \text{Model count}
  \end{tikzcd}
\end{frame}

\begin{frame}[fragile]{Finding (a Sufficient Set of) Base Cases}
  \begin{block}{Outline}
    \begin{enumerate}
      \item For every function call:
            \begin{enumerate}
              \item For every argument of the form
                    \structure{$\textit{var} - \textit{const}$}:
                    \begin{enumerate}
                      \item Replace the signature parameter with
                            \structure{$0$}, \structure{$1$}, \dots,
                            \structure{$\textit{const} - 1$}
                    \end{enumerate}
              \item For every argument of the form \structure{$\textit{const}$}:
                    \begin{enumerate}
                      \item Replace the corresponding signature parameter with
                            \structure{$\textit{const}$}
                    \end{enumerate}
            \end{enumerate}
    \end{enumerate}
  \end{block}
  \begin{example}
    The signature of \structure{$g$} is \structure{$g(l, m)$}.

@@

% TODO: propagate + maybe smoothing
\begin{frame}{From a Base Case to a Sentence}
  \begin{block}{Preliminaries}
    \begin{itemize}
      \item \textsc{Crane} associates each function \structure{$f$} with a
            sentence \structure{$\phi$} such that
            \structure{$\textsc{Crane}(\phi)$} produces the definition of
            \structure{$f$}
      \item There is a bijection between the parameters of \structure{$f$} and
            the domains of \structure{$\phi$}
    \end{itemize}
  \end{block}
\end{frame}

>>>>

\documentclass{beamer}
\usepackage[T1]{fontenc}
\usepackage[utf8]{inputenc}
\usepackage[UKenglish]{babel}
\usepackage[UKenglish]{isodate}
\usepackage{tikz}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{complexity}
\usepackage{fontawesome}
\usepackage{dashbox}
\usepackage[style=authoryear]{biblatex}

\addbibresource{talk.bib}

@@

  \begin{tikzcd}
    g(0, m) \arrow[d, gray, visible on=<2->] & \phantom{-} \arrow[d, phantom, visible on=<2->, "\text{\structure{4.1. ($\Rightarrow$)} Construct the corr.\ sentence}"] & g(l, 0) \arrow[d, gray, visible on=<2->] \\
    |[visible on=<2->]| \forall y \in \Delta\text{. }S(y) \lor \neg S(y) \arrow[d, gray, visible on=<3->] & \phantom{-} \arrow[d, phantom, visible on=<3->, "\text{\structure{4.2.} Recurse}"] & |[visible on=<2->]| \top \arrow[d, gray, visible on=<3->] \\
    |[visible on=<3->]| g(0, m) = \alt<4->{0^m}{\text{???}} \arrow[rd, gray, visible on=<5->] & |[visible on=<5->]| \text{\structure{4.3.} Add to }\structure{\mathcal{E}} & |[visible on=<3->]| g(l, 0) = \alt<4->{1}{???} \arrow[ld, gray, visible on=<5->] \\
    & |[visible on=<5->]| \mathcal{E} \arrow[d, gray, visible on=<6->, "\text{\color{black}{\structure{5.} Compile to C++}}"] & \\
    |[visible on=<7->]| \text{Domain sizes} \arrow[r, gray, visible on=<7->] & |[visible on=<6->]| \text{\dbox{C++ code}} \arrow[r, gray, visible on=<7->] & |[visible on=<7->]| \text{Model count}
  \end{tikzcd}
\end{frame}

\begin{frame}[fragile]{Finding (a Sufficient Set of) Base Cases}
  \begin{block}{Outline}
    \begin{enumerate}
      \item For every \alert{function call}:
            \begin{enumerate}
              \item For every \alert{argument} of the form
                    \structure{$\textit{var} - \textit{const}$}:
                    \begin{enumerate}
                      \item Replace the \alert{signature parameter} with
                            \structure{$0$}, \structure{$1$}, \dots,
                            \structure{$\textit{const} - 1$}
                    \end{enumerate}
              \item For every \alert{argument} of the form
                    \structure{$\textit{const}$}:
                    \begin{enumerate}
                      \item Replace the corresponding signature parameter with
                            \structure{$\textit{const}$}
                    \end{enumerate}
            \end{enumerate}
    \end{enumerate}
  \end{block}
  \begin{example}
    The \alert{signature} of \structure{$g$} is \structure{$g(l, m)$}.

@@

\begin{frame}[fragile]{From a Base Case to a Sentence}
  \begin{block}{From Previous Work~\color{gray}{\parencite{DBLP:conf/kr/DilkasB23}}}
    \begin{itemize}
      \item \textsc{Crane} associates each function \structure{$f$} with a
            sentence \structure{$\phi$} such that
            \structure{$\textsc{Crane}(\phi)$} produces the definition of
            \structure{$f$}
      \item There is a bijection between the parameters of \structure{$f$} and
            the domains of \structure{$\phi$}
    \end{itemize}
  \end{block}
  \begin{example}
    \begin{itemize}
      \item Base case: \structure{$g(\stackrel{\mbox{\color{black}{\tiny\onslide<2->{\ensuremath{\Gamma}}}}}{\stackrel{\mbox{\color{black}{\tiny\onslide<2->{\ensuremath{\updownarrow}}}}}{0}}, \stackrel{\mbox{\color{black}{\tiny\onslide<2->{\ensuremath{\Delta}}}}}{\stackrel{\mbox{\color{black}{\tiny\onslide<2->{\ensuremath{\updownarrow}}}}}{m}})$}
      \item<3-> \alert{Part} of the sentence of \structure{$g$}:
            \begin{equation}\label{eq}
              \forall x \in \Gamma\text{. }\forall y \in \Delta\text{. }S(y) \lor \neg P(x, y)
            \end{equation}
      \item<4-> \structure{$g(0, \dots)$} means we need to simplify \eqref{eq}
            by assuming \structure{$|\Gamma| = 0$}
      \item<5-> Result: \structure{$\forall y \in \Delta\text{.
            }S(y) \lor \neg S(y)$} \qquad (\alert{Smoothing})
    \end{itemize}
  \end{example}
\end{frame}
